---
title: "Survival analysis renal"
subtitle: <center> Survival Analysis</center>
author: "Mario Presti"
date: '`r paste("First created on Feb 2025. Updated on ", format(Sys.Date(), "%d %B %Y"))`'
output:
  html_document:
    css: style.css
    code_folding: hide
    fig_caption: yes
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 5
    number_sections: true
---

# Introduction

Multivariate Survival Analysis on the Renal database.

# Loading Libraries

```{r setup,warning=F, message=F,eval=TRUE}
# Load required libraries
library(pacman)
pacman::p_load(data.table,stringr,tidyverse,ggplot2,MatchIt,survival,survminer,survey,compareGroups, forestmodel, forestplot, openxlsx,kableExtra,ggsurvfit,extrafont,DT,tibble,strex,hrbrthemes,ggstatsplot,reshape,pander,ggrepel,scales,dataMaid,gridExtra,tidytidbits,survivalAnalysis,gtsummary, Cairo,Amelia,officer,mice,naniar)
```
## Pre-processing of data

* Check that all columns have no spaces etc, especially character columns, 
* transform character columns to factors. 
* Check if all factors are ok.
* Check numerical columns.

We first need to define which columns we want as character (categorical), and which need to be numerical. Sometimes things get messed up with excel files.

```{r Data loading and correction,warning=F, message=F,eval=TRUE}
setwd("E:/PhD_projects/Realworld/Data/")
dataDF <- read.xlsx("renal.xlsx")

# Check for duplicates
any(duplicated(dataDF$patient_id))

# Remove duplicates
dataDF <- dataDF[!duplicated(dataDF), ]
dim(dataDF)

# Rename some columns
names(dataDF)[names(dataDF) == 'progression_Marco'] <- 'Progressed'
names(dataDF)[names(dataDF) == 'age'] <- 'age_1st_treat'
names(dataDF)[names(dataDF) == 'Treatment_line'] <- 'line_correct'
names(dataDF)[names(dataDF) == 'Treatment'] <- 'regime_correct'
names(dataDF)[names(dataDF) == 'CNS_mets'] <- 'Brain_metastases'
names(dataDF)[names(dataDF) == 'BOR'] <- 'bor'
names(dataDF)[names(dataDF) == 'ecog_ps.1'] <- 'PS'

# Define the features of interest
features <- c("sex", "age_1st_treat","line_correct", "regime_correct", 
              "Brain_metastases","PS", "bor")

features_cancer_spec <- c("clear_cell","sarcomatoid", "smoking", "IMDC")

## Define which columns include survival data
OS_time <- c("OS_days")
OS_status <- c("Dead")
PFS_time <- c("PFS_days")
PFS_status <- c("Progressed")

## DEFINE WHICH COLUMNS ARE CATEGORICAL
categ_feats <-c("sex","regime_correct","PS", "bor", "line_correct",features_cancer_spec)

## DEFINE WHICH COLUMNS ARE NUMERICAL
numeric_feats <- c("OS_days", "Dead","PFS_days","Progressed","age_1st_treat")

# Remove spaces from column names and features.
colnames(dataDF) <- gsub(" ","_",colnames(dataDF))
features <- gsub(" ","_",features)
categ_feats <- gsub(" ","_",categ_feats)
numeric_feats <- gsub(" ","_",numeric_feats)
OS_time <- gsub(" ","_",OS_time)
OS_status <- gsub(" ","_",OS_status)
PFS_time <- gsub(" ","_",PFS_time)
PFS_status <- gsub(" ","_",PFS_status)

## MAKE SURE THERE ARE NO BLANKS ANYWHERE IN THE DATA
## Find non-Numeric columns
num_cols <- sapply(dataDF, is.numeric)

dataDF <- cbind(apply(dataDF[,!num_cols], 2, str_remove_all, "NA"),dataDF[,num_cols])   # Remove blanks

### MAKE SURE THAT THE COLUMNS ARE IN THE CORRECT FORMAT NEEDED FOR ANALYSIS.
## Numeric columns
dataDF <- dataDF %>% mutate(across(all_of(numeric_feats), as.numeric))

## Categorical
dataDF <- dataDF %>% mutate(across(all_of(categ_feats), as.character))

# Transform character columns to categorical, to have the different levels
dataDF <- dataDF %>% dplyr::mutate(across(all_of(categ_feats), as.factor))
dataDF <- dataDF %>% dplyr::select(all_of(c(colnames(dataDF)[1], OS_time,OS_status, PFS_time,PFS_status, features, features_cancer_spec))) %>% distinct()


print("Numbers per treatment")
table(dataDF$regime_correct)

print("Numbers per response")
table(dataDF$bor)

## SUMMARY OF THE DATA
summary(dataDF)

# remove some white spaces in the PS column
dataDF$PS <- trimws(dataDF$PS)
dataDF$PS <- as.factor(dataDF$PS)

# Rename some factors for plots
dataDF <- dataDF %>%
  dplyr::mutate(
    Sex = case_when(
      sex == "2"   ~ "Female",
      sex == "1"   ~ "Male",
      sex == "99"  ~ NA_character_,
      TRUE         ~ NA_character_
    ),
    Sex = factor(Sex, levels = c("Male", "Female")),

    regime_correct = case_when(
      regime_correct == "nivolumab" ~ "Anti-PD1",
      regime_correct == "ipi/nivo"  ~ "Ipi+Nivo",
      regime_correct == "99"        ~ NA_character_,
      TRUE                           ~ NA_character_
    ),
    regime_correct = factor(
      regime_correct,
      levels = c("Anti-PD1", "Ipi+Nivo")
    ),

    PS = case_when(
      PS %in% c("1", "2", "3") ~ "PS≥1",
      PS == "0"               ~ "PS=0",
      PS == "99"              ~ NA_character_,
      TRUE                    ~ NA_character_
    ),
    PS = factor(PS, levels = c("PS=0", "PS≥1")),
    
    regime_correct = case_when(
      regime_correct == "Ipi+Nivo" ~ "Anti-PD1+Anti-CTLA4",
      regime_correct == "Anti-PD1" ~ "Anti-PD1",
      regime_correct == "99"       ~ NA_character_,
      TRUE                    ~ NA_character_
    ),
    regime_correct = factor(regime_correct, levels = c("Anti-PD1", "Anti-PD1+Anti-CTLA4")),

    smoking = case_when(
      smoking == "0"           ~ "Non smoker",
      smoking == "1" ~ "Current smoker",
      smoking == "2" ~ "Former smoker",
      smoking == "99"          ~ NA_character_,
      TRUE                     ~ NA_character_
    ),
    smoking = factor(
      smoking,
      levels = c("Non smoker", "Current smoker","Former smoker")
    ),
    
    Brain_metastases = case_when(
     Brain_metastases == "No"           ~ "No",
     Brain_metastases == "Yes" ~ "Yes",
      TRUE                     ~ NA_character_
    ),
    Brain_metastases = factor(
      Brain_metastases,
      levels = c("No", "Yes")
    ),

    clear_cell = case_when(
      clear_cell == "ja"   ~ "Clear cell",
      clear_cell == "nej"  ~ "Non clear cell",
      clear_cell == "99"   ~ NA_character_,
      TRUE                  ~ NA_character_
    ),
    clear_cell = factor(
      clear_cell,
      levels = c("Clear cell", "Non clear cell")
    ),
    sarcomatoid = case_when(
      sarcomatoid == "ja"   ~ "Sarcomatoid",
      sarcomatoid == "nej"  ~ "Non sarcomatoid",
      sarcomatoid == "99"   ~ NA_character_,
      TRUE                  ~ NA_character_
    ),
    sarcomatoid = factor(
      sarcomatoid,
      levels = c("Sarcomatoid", "Non sarcomatoid")
    ),
    IMDC = case_when(
      IMDC == "Poor"   ~ "Poor",
      IMDC == "Good/Intermediate"  ~ "Good/Intermediate",
      IMDC == "99"   ~ NA_character_,
      TRUE                  ~ NA_character_
    ),
    IMDC = factor(
      IMDC,
      levels = c("Poor", "Good/Intermediate")
    )
  )

#remove some useless columns
dataDF$sex <- NULL

# rename some of the columns for plotting
names(dataDF)[names(dataDF) == 'age_1st_treat'] <- 'Age'
names(dataDF)[names(dataDF) == 'regime_correct'] <- 'CPI Regimen'
names(dataDF)[names(dataDF) == 'PS'] <- 'ECOG Performance Status'
# names(dataDF)[names(dataDF) == 'stage_2'] <- 'AJCC 8th stage'
names(dataDF)[names(dataDF) == 'Brain_metastases'] <- 'Brain metastases'
# names(dataDF)[names(dataDF) == 'braf_correct'] <- 'BRAF mutation'
names(dataDF)[names(dataDF) == 'bor'] <- 'Objective response'
names(dataDF)[names(dataDF) == 'line_correct'] <- 'Treatment line'
names(dataDF)[names(dataDF) == 'clear_cell'] <- 'RCC subtype'
names(dataDF)[names(dataDF) == 'sarcomatoid'] <- 'Sarcomatoid subtype'
names(dataDF)[names(dataDF) == 'smoking'] <- 'Smoking status'
features <- c("Sex", "Age", "CPI Regimen", "Treatment line", "RCC subtype", "Sarcomatoid subtype","IMDC","ECOG Performance Status", "Objective response")

write.csv(dataDF, file = "RCC_data_polished.csv", sep = ",", append = F, row.names = F)

#Administrative censoring at 60 months
dataDF$censor_time_OS   <- pmin(dataDF$OS_days/30, 60)
dataDF$censor_status_OS <- ifelse(dataDF$OS_days/30 > 60, 0, dataDF$Dead)
#Administrative censoring at 60 months
dataDF$censor_time_PFS   <- pmin(dataDF$PFS_days/30, 60)
dataDF$censor_status_PFS <- ifelse(dataDF$PFS_days/30 > 60, 0, dataDF$Progressed)
```

```{r select descriptive features,warning=F, message=F,eval=TRUE}
descr_features <- paste0("`", c(features, "CPI Regimen"), "`", collapse = " + ")
descr_formula <- as.formula(paste0("~", descr_features))

Baseline_Characteristics_table_Partial_patients <- descrTable(descr_formula , data = dataDF)
```

# OS analysis
##Multiple Univariate analysis - Including missing data
```{r Multiple Univariate analysis - Including missing data}
surv_time <- "censor_time_OS"
surv_status <- "censor_status_OS"
surv_time_label <- "OS"
# Find the best reference for each feature
best_refs <- character(length(features))
names(best_refs) <- features

for (i in seq_along(features)) {
  var <- features[i]

  # Skip this iteration if this column isn't a factor
  if (! is.factor(dataDF[[var]])) {
    next
  }

  vals <- na.omit(unique(dataDF[[var]]))
  # ensure factor
  vals <- as.character(vals)

  # storage for max HR per candidate ref
  hr_max <- numeric(length(vals))

  for (j in seq_along(vals)) {
    ref_level <- vals[j]
    df2 <- dataDF
    if(!"ordered" %in% class(df2[[var]])){
      df2[[var]] <- relevel(as.factor(df2[[var]]), ref = ref_level)
    }

    uni <- analyse_survival(
      df2,
      time_status = c(surv_time, surv_status),
      by          = .data[[var]]
    )

    fit   <- uni$coxph    # the coxph object
    coefs <- coef(fit)    # log HRs
    hrs   <- exp(coefs)   # HRs
    hr_max[j] <- if (length(hrs) > 0) max(hrs, na.rm = TRUE) else NA
  }

  # pick the ref that yielded the largest HR
  best_refs[var] <- vals[which.max(hr_max)]
}

# Re‐level dataDF based on those best refs
for (var in features) {
  if (! is.factor(dataDF[[var]])) {
    next
  }
  if(!"ordered" %in% class(dataDF[[var]])){
    dataDF[[var]] <- relevel(
    as.factor(dataDF[[var]]),
    ref = best_refs[var]
  )
  }
}

multiple_uni <- map(features, function(by){analyse_multivariate(dataDF,
                       c(surv_time,surv_status),
                       covariates = list(by), # covariates expects a list
                       covariate_name_dict = NULL,
                       covariate_label_dict = NULL)})
max_cis <- max(sapply(
  multiple_uni,
  function(x) max(x$summary$conf.int, na.rm = TRUE))
)
min_cis <- min(sapply(
  multiple_uni,
  function(x) min(x$summary$conf.int, na.rm = TRUE))
)

HR_x_limits <- c(min_cis, max_cis) * c(1/1.5, 1.5)
HR_x_breaks <- round(10^pretty(log10(HR_x_limits), n = 7),digits = 1)

forest_plot(multiple_uni,
            #factor_labeller = covariate_names,
            endpoint_labeller = c(time=surv_time_label),
            # orderer = ~order(HR),
            labels_displayed = c("factor"),#"endpoint",
            ggtheme = ggplot2::theme_bw(base_size = 10),
            #values_displayed = c("HR","CI","p"),
            HR_x_limits = c(min_cis*1.5,max_cis*1.5),
            HR_x_breaks = c(1,2,ceiling(10^pretty(log10(HR_x_limits), n = 7)*2/2))
            #p_lessthan_cutoff = 0.05
            )
```
## Multivariate Survival Analysis - only complete cases

```{r  Multivariate Survival Analysis - only complete cases, fig.width=10, fig.height=8, warning=FALSE, message=FALSE, eval=TRUE}
# # To remove both (NAs and empty):
dataDF_complete <- dataDF %>%
  drop_na() %>%  # Remove NA values
  filter_all(all_vars(. != "")) %>%  # Remove empty strings
  filter_all(all_vars(trimws(.) != ""))  # Remove strings with only spaces

dim(dataDF_complete)

multiv_feats_sign <- c()
for (i in seq_along(features)){
  coef <- as.data.frame(multiple_uni[[i]]$summary$coefficients)
  if(nrow(coef) >1){
    pval <- min(coef$`Pr(>|z|)`)
  } else {
  pval <- coef$`Pr(>|z|)`
  }
  if(any(pval < 0.2)){
    multiv_feats_sign <- c(multiv_feats_sign, multiple_uni[[i]]$overall$covariates)
  }
}

multiv_feats_sign_formula <- paste0("`", multiv_feats_sign, "`", collapse = " + ")
fmla_sign <- as.formula(paste0(
  "Surv(dataDF_complete[[surv_time]], dataDF_complete[[surv_status]]) ~ ",
  multiv_feats_sign_formula
))

multiv_feats_all<- paste0("`", features, "`", collapse = " + ")
fmla_all <- as.formula(paste0(
  "Surv(dataDF_complete[[surv_time]], dataDF_complete[[surv_status]]) ~ ",
  multiv_feats_all
))

cox_model_all <- coxph(fmla_all, data = dataDF_complete)
cox_model_sign <- coxph(fmla_sign, data = dataDF_complete)

summary(cox_model_all)
summary(cox_model_sign)

forest_model_all <- forest_model(cox_model_all,  # Set x-axis limits
                             exponentiate = TRUE)  # Display hazard ratios on log scale)
forest_model_sign <- forest_model(cox_model_sign,  # Set x-axis limits
                             exponentiate = TRUE)  # Display hazard ratios on log scale)

# Display the plot
print(forest_model_all)
print(forest_model_sign)
```

# Combined Univariable vs Multivariable Forest Plots - OS
```{r Combined Univariable vs Multivariable Forest Plots - OS, fig.width=10, fig.height=7, warning=FALSE, message=FALSE, eval=TRUE}
# Prepare empty list to collect rows
rows <- vector("list", 0)

# Summaries of the multivariable model
s_multi    <- summary(cox_model_sign)
coef_multi <- s_multi$coefficients      # matrix with exp(coef) and p
ci_multi   <- s_multi$conf.int          # matrix with lower/upper .95

# Loop over each univariable result
for (i in seq_along(multiple_uni)) {
  u    <- multiple_uni[[i]]
  cov  <- u$overall$covariates           # the covariate name
  fit  <- u$coxph
  s_u  <- summary(fit)
  coef_u <- s_u$coefficients             # matrix with exp(coef) etc.
  ci_u   <- s_u$conf.int
  if(u$summaryAsFrame$factor.value[1] == "<continuous>"){
    reference <- "NA"
    levs <- "Fitted as continuous"
  } else {
    reference <- levels(as.factor(u$data[,ncol(u$data)]))[1]
    levs <- setdiff(unique(u$data[,ncol(u$data)]), reference)
  }
  # for each row (level) in the univariable fit:
  for (j in seq_len(nrow(coef_u))) {
    HRu  <- coef_u[j, "exp(coef)"]
    LUu  <- ci_u[j,   "lower .95"]
    UCu  <- ci_u[j,   "upper .95"]
    pu   <- coef_u[j, "Pr(>|z|)"]
    
    if (cov %in% multiv_feats_sign) {
      HRm <- coef_multi[grep(cov, rownames(coef_multi))[j], "exp(coef)"]
      LUm <- ci_multi[ grep(cov, rownames(coef_multi))[j], "lower .95"]
      UCm <- ci_multi[ grep(cov, rownames(coef_multi))[j], "upper .95"]
      pm  <- coef_multi[grep(cov, rownames(coef_multi))[j], "Pr(>|z|)"]
    } else {
      HRm <- NA; LUm <- NA; UCm <- NA; pm <- NA
    }
    
    rows[[length(rows) + 1]] <- data.frame(
      Covariate       = cov,
      Level           = levs[j],
      Reference       = reference,
      HR_uni          = HRu,
      CI_lo_uni       = LUu,
      CI_hi_uni       = UCu,
      p_uni           = pu,
      HR_multi        = HRm,
      CI_lo_multi     = LUm,
      CI_hi_multi     = UCm,
      p_multi         = pm,
      stringsAsFactors = FALSE
    )
  }
}

# Bind into one data.frame
results_df <- do.call(rbind, rows)

# Format up to three digits
results_df <- results_df %>%
  mutate(across(starts_with("HR_") | starts_with("CI_") | matches("^p_"),
                ~ signif(.x, 3)))

results_df$CI_lo_uni <- paste0(results_df$CI_lo_uni, "-", results_df$CI_hi_uni)
results_df$CI_hi_uni <- NULL
results_df$CI_lo_multi <- paste0(results_df$CI_lo_multi, "-", results_df$CI_hi_multi)
results_df$CI_hi_multi <- NULL
colnames(results_df) <- c("Covariate", "Level", "Reference" , "Univariable HR", "Univariable CI", "Univariable P-value", "Multivariable HR", "Multivariable CI", "Multivariable P-value")

last_col <- ncol(results_df)
results_df <- results_df %>%
  mutate(across(
    .cols = -last_col,
    .fns  = ~ ifelse(is.na(.) | grepl("NA", ., fixed = TRUE), "", .)
  )) %>%
  mutate(across(
    .cols = last_col,
    .fns  = ~ ifelse(is.na(.), "Excluded from multivariable analysis", as.character(.))
  ))

results_df <- results_df %>%
  group_by(Covariate) %>%
  mutate(Covariate = ifelse(row_number() == 1, Covariate, "")) %>%
  ungroup()

n_cols <- ncol(results_df)

# align: left for first two, center for the rest
aligns <- c("l", "l", rep("c", n_cols - 2))

results_df %>%
  kable(
    format = "html",
    caption = "Univariable vs Multivariable Cox Results - OS",
    align   = aligns,
    escape  = FALSE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  # make header row bold and centered
  row_spec(
    0,
    bold  = TRUE,
    align = "center"
  ) %>%
  # add vertical lines right of col 3 and col 6
  column_spec(3, border_right = TRUE) %>%
  column_spec(6, border_right = TRUE)
```


# PFS analysis
## Multiple Univariate analysis - Incomplete complete cases
```{r Multiple Univariate analysis - PFS}
surv_time <- "censor_time_PFS"
surv_status <- "censor_status_PFS"
surv_time_label <- "PFS"
# Find the best reference for each feature
best_refs <- character(length(features))
names(best_refs) <- features

for (i in seq_along(features)) {
  var <- features[i]

  # Skip this iteration if this column isn't a factor
  if (! is.factor(dataDF[[var]])) {
    next
  }

  vals <- na.omit(unique(dataDF[[var]]))
  # ensure factor
  vals <- as.character(vals)

  # storage for max HR per candidate ref
  hr_max <- numeric(length(vals))

  for (j in seq_along(vals)) {
    ref_level <- vals[j]
    df2 <- dataDF
    if(!"ordered" %in% class(df2[[var]])){
      df2[[var]] <- relevel(as.factor(df2[[var]]), ref = ref_level)
    }

    uni <- analyse_survival(
      df2,
      time_status = c(surv_time, surv_status),
      by          = .data[[var]]
    )

    fit   <- uni$coxph    # the coxph object
    coefs <- coef(fit)    # log HRs
    hrs   <- exp(coefs)   # HRs
    hr_max[j] <- if (length(hrs) > 0) max(hrs, na.rm = TRUE) else NA
  }

  # pick the ref that yielded the largest HR
  best_refs[var] <- vals[which.max(hr_max)]
}

# Re‐level dataDF based on those best refs
for (var in features) {
  if (! is.factor(dataDF[[var]])) {
    next
  }
  if(!"ordered" %in% class(dataDF[[var]])){
    dataDF[[var]] <- relevel(
    as.factor(dataDF[[var]]),
    ref = best_refs[var]
  )
  }
}

multiple_uni <- map(features, function(by){analyse_multivariate(dataDF,
                       c(surv_time,surv_status),
                       covariates = list(by), # covariates expects a list
                       covariate_name_dict = NULL,
                       covariate_label_dict = NULL)})
max_cis <- max(sapply(
  multiple_uni,
  function(x) max(x$summary$conf.int, na.rm = TRUE))
)
min_cis <- min(sapply(
  multiple_uni,
  function(x) min(x$summary$conf.int, na.rm = TRUE))
)

HR_x_limits <- c(min_cis, max_cis) * c(1/1.5, 1.5)
HR_x_breaks <- round(10^pretty(log10(HR_x_limits), n = 7),digits = 1)

forest_plot(multiple_uni,
            #factor_labeller = covariate_names,
            endpoint_labeller = c(time=surv_time_label),
            # orderer = ~order(HR),
            labels_displayed = c("factor"),#"endpoint",
            ggtheme = ggplot2::theme_bw(base_size = 10),
            #values_displayed = c("HR","CI","p"),
            HR_x_limits = c(min_cis*1.5,max_cis*1.5),
            HR_x_breaks = c(1,2,ceiling(10^pretty(log10(HR_x_limits), n = 7)*2/2))
            #p_lessthan_cutoff = 0.05
            )
```

# Multivariate Survival Analysis
## All responders - PFS
```{r, fig.width=10, fig.height=8, warning=FALSE, message=FALSE, eval=TRUE}
multiv_feats_sign <- c()
for (i in seq_along(features)){
  coef <- as.data.frame(multiple_uni[[i]]$summary$coefficients)
  if(nrow(coef) >1){
    pval <- min(coef$`Pr(>|z|)`)
  } else {
  pval <- coef$`Pr(>|z|)`
  }
  if(any(pval < 0.2)){
    multiv_feats_sign <- c(multiv_feats_sign, multiple_uni[[i]]$overall$covariates)
  }
}

multiv_feats_sign_formula <- paste0("`", multiv_feats_sign, "`", collapse = " + ")
fmla_sign <- as.formula(paste0(
  "Surv(dataDF_complete[[surv_time]], dataDF_complete[[surv_status]]) ~ ",
  multiv_feats_sign_formula
))

multiv_feats_all<- paste0("`", features, "`", collapse = " + ")
fmla_all <- as.formula(paste0(
  "Surv(dataDF_complete[[surv_time]], dataDF_complete[[surv_status]]) ~ ",
  multiv_feats_all
))

cox_model_all <- coxph(fmla_all, data = dataDF_complete)
cox_model_sign <- coxph(fmla_sign, data = dataDF_complete)

summary(cox_model_all)
summary(cox_model_sign)

forest_model_all <- forest_model(cox_model_all,  # Set x-axis limits
                             exponentiate = TRUE)  # Display hazard ratios on log scale)
forest_model_sign <- forest_model(cox_model_sign,  # Set x-axis limits
                             exponentiate = TRUE)  # Display hazard ratios on log scale)

# Display the plot
print(forest_model_all)
print(forest_model_sign)

```

# Combined Univariable vs Multivariable Forest Plots - PFS
```{r Combined Univariable vs Multivariable Forest Plots - PFS, fig.width=10, fig.height=7, warning=FALSE, message=FALSE, eval=TRUE}
# Prepare empty list to collect rows
rows <- vector("list", 0)

# Summaries of the multivariable model
s_multi    <- summary(cox_model_sign)
coef_multi <- s_multi$coefficients      # matrix with exp(coef) and p
ci_multi   <- s_multi$conf.int          # matrix with lower/upper .95

# Loop over each univariable result
for (i in seq_along(multiple_uni)) {
  u    <- multiple_uni[[i]]
  cov  <- u$overall$covariates           # the covariate name
  fit  <- u$coxph
  s_u  <- summary(fit)
  coef_u <- s_u$coefficients             # matrix with exp(coef) etc.
  ci_u   <- s_u$conf.int
  if(u$summaryAsFrame$factor.value[1] == "<continuous>"){
    reference <- "NA"
    levs <- "Fitted as continuous"
  } else {
    reference <- levels(as.factor(u$data[,ncol(u$data)]))[1]
    levs <- setdiff(unique(u$data[,ncol(u$data)]), reference)
  }
  # for each row (level) in the univariable fit:
  for (j in seq_len(nrow(coef_u))) {
    HRu  <- coef_u[j, "exp(coef)"]
    LUu  <- ci_u[j,   "lower .95"]
    UCu  <- ci_u[j,   "upper .95"]
    pu   <- coef_u[j, "Pr(>|z|)"]
    
    if (cov %in% multiv_feats_sign) {
      HRm <- coef_multi[grep(cov, rownames(coef_multi))[j], "exp(coef)"]
      LUm <- ci_multi[ grep(cov, rownames(coef_multi))[j], "lower .95"]
      UCm <- ci_multi[ grep(cov, rownames(coef_multi))[j], "upper .95"]
      pm  <- coef_multi[grep(cov, rownames(coef_multi))[j], "Pr(>|z|)"]
    } else {
      HRm <- NA; LUm <- NA; UCm <- NA; pm <- NA
    }
    
    rows[[length(rows) + 1]] <- data.frame(
      Covariate       = cov,
      Level           = levs[j],
      Reference       = reference,
      HR_uni          = HRu,
      CI_lo_uni       = LUu,
      CI_hi_uni       = UCu,
      p_uni           = pu,
      HR_multi        = HRm,
      CI_lo_multi     = LUm,
      CI_hi_multi     = UCm,
      p_multi         = pm,
      stringsAsFactors = FALSE
    )
  }
}

# Bind into one data.frame
results_df <- do.call(rbind, rows)

# Format up to three digits 
results_df <- results_df %>%
  mutate(across(starts_with("HR_") | starts_with("CI_") | matches("^p_"),
                ~ signif(.x, 3)))

results_df$CI_lo_uni <- paste0(results_df$CI_lo_uni, "-", results_df$CI_hi_uni)
results_df$CI_hi_uni <- NULL
results_df$CI_lo_multi <- paste0(results_df$CI_lo_multi, "-", results_df$CI_hi_multi)
results_df$CI_hi_multi <- NULL
colnames(results_df) <- c("Covariate", "Level", "Reference" , "Univariable HR", "Univariable CI", "Univariable P-value", "Multivariable HR", "Multivariable CI", "Multivariable P-value")

last_col <- ncol(results_df)
results_df <- results_df %>%
  mutate(across(
    .cols = -last_col,
    .fns  = ~ ifelse(is.na(.) | grepl("NA", ., fixed = TRUE), "", .)
  )) %>%
  mutate(across(
    .cols = last_col,
    .fns  = ~ ifelse(is.na(.), "Excluded from multivariable analysis", as.character(.))
  ))

results_df$`Univariable P-value` <- round(results_df$`Univariable P-value`, digits = 3)
# Now print
knitr::kable(
  results_df,
  caption = "Univariable vs Multivariable Cox Results",
  row.names = FALSE
)

n_cols <- ncol(results_df)

# align: left for first two, center for the rest
aligns <- c("l", "l", rep("c", n_cols - 2))

results_df %>%
  kable(
    format = "html",
    caption = "Univariable vs Multivariable Cox Results - PFS",
    align   = aligns,
    escape  = FALSE
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  # make header row bold and centered
  row_spec(
    0,
    bold  = TRUE,
    align = "center"
  ) %>%
  # add vertical lines right of col 3 and col 6
  column_spec(3, border_right = TRUE) %>%
  column_spec(6, border_right = TRUE)

wb <- createWorkbook()
addWorksheet(wb, "Supp_Table_Cox")

# Write data
writeData(wb, sheet = "Supp_Table_Cox", x = results_df, startRow = 1, startCol = 1)

# Compute dimensions
n_rows <- nrow(results_df) + 1  # +1 for header
n_cols <- ncol(results_df)

# Define styles
headerStyle <- createStyle(
  fontSize        = 12,
  textDecoration  = "bold",
  halign          = "center",
  border          = "bottom",
  borderStyle     = "medium"
)
evenRowStyle <- createStyle(fgFill = "#F2F2F2")
sepBorderStyle <- createStyle(border = "right", borderStyle = "thin")

# Apply alternating shading
dataRows <- 2:n_rows
evenRows <- dataRows[seq(1, length(dataRows), by = 2)]

# Add vertical separators after col 3 and 6
addStyle(
  wb, "Supp_Table_Cox",
  style = sepBorderStyle,
  rows  = 1:n_rows,
  cols  = 3,
  gridExpand = TRUE
)
addStyle(
  wb, "Supp_Table_Cox",
  style = sepBorderStyle,
  rows  = 1:n_rows,
  cols  = 6,
  gridExpand = TRUE
)

#alternate shading
addStyle(
  wb, "Supp_Table_Cox",
  style = evenRowStyle,
  rows  = evenRows,
  cols  = 1:n_cols,
  gridExpand = TRUE
)

# Apply header style
addStyle(
  wb, "Supp_Table_Cox",
  style = headerStyle,
  rows  = 1, cols = 1:n_cols,
  gridExpand = TRUE
)

# 7) Autofit column widths & freeze header
setColWidths(wb, "Supp_Table_Cox", cols = 1:n_cols, widths = "auto")
freezePane(wb, "Supp_Table_Cox", firstRow = TRUE)

# 8) Save file
saveWorkbook(
  wb,
  file = "E:/PhD_projects/Realworld/Scripts/Acquired_resistance_RW/Tables/Supplementary_Table_3.xlsx",
  overwrite = TRUE
)
```


# Five-Year Survival Estimates by Response

```{r five_year_survival_CI, warning=FALSE, message=FALSE}
library(survival)
library(dplyr)

# the two response groups
responses <- c("CR", "PR")

# helper to get survival ± CI at 60 months
get5yr_ci <- function(fit) {
  s <- summary(fit, times = 60)
  if (length(s$surv) == 0) {
    s_all   <- summary(fit)
    surv    <- tail(s_all$surv, 1)
    lower   <- tail(s_all$lower, 1)
    upper   <- tail(s_all$upper, 1)
  } else {
    surv  <- s$surv
    lower <- s$lower
    upper <- s$upper
  }
  data.frame(
    Estimate = surv,
    Lower95  = lower,
    Upper95  = upper
  )
}

# loop over response levels
results <- lapply(responses, function(resp) {
  d <- filter(dataDF, `Objective response` == resp)
  fit_OS  <- survfit(Surv(censor_time_OS,  censor_status_OS)  ~ 1, data = d)
  fit_PFS <- survfit(Surv(censor_time_PFS, censor_status_PFS) ~ 1, data = d)

  os_ci  <- get5yr_ci(fit_OS)
  pfs_ci <- get5yr_ci(fit_PFS)

  data.frame(
    Response       = resp,
    OS_5yr         = os_ci$Estimate,
    OS_5yr_Lower   = os_ci$Lower95,
    OS_5yr_Upper   = os_ci$Upper95,
    PFS_5yr        = pfs_ci$Estimate,
    PFS_5yr_Lower  = pfs_ci$Lower95,
    PFS_5yr_Upper  = pfs_ci$Upper95
  )
})

five_year_table <- bind_rows(results) %>%
  mutate(across(matches("_5yr"), ~ . * 100, .names = "{.col}_pct")) %>%
  select(Response,
         OS_5yr_pct, OS_5yr_Lower_pct, OS_5yr_Upper_pct,
         PFS_5yr_pct, PFS_5yr_Lower_pct, PFS_5yr_Upper_pct)
print(five_year_table)
```

```{r - 5 year KM curves, fig.width=10, fig.height=8}
# pull out labels from five_year_table
x_max=60
rownames(five_year_table) <- five_year_table$Response
os_lbls <- five_year_table %>%
  transmute(
    Response,
    os_lbls = sprintf("%s: %.0f%% (%.0f–%.0f%%)",
                       Response, OS_5yr_pct, OS_5yr_Lower_pct, OS_5yr_Upper_pct))

pfs_lbls <- five_year_table %>%
  transmute(
    Response,
    pfs_lbls = sprintf("%s: %.0f%% (%.0f–%.0f%%)",
                        Response, PFS_5yr_pct, PFS_5yr_Lower_pct, PFS_5yr_Upper_pct))

#adding a syntactically correct name for OR, as it was failing in the Survfit step
dataDF$Objective_response = dataDF$`Objective response`

ypos <- c(CR = five_year_table$OS_5yr_pct[1], PR = five_year_table$OS_5yr_pct[2])

fit_OS <- survfit(
  Surv(censor_time_OS, censor_status_OS) ~ Objective_response,
  data = dataDF
)
names(fit_OS$strata) <- gsub("Objective_response=", "", names(fit_OS$strata))
# OS plot
p_os <- ggsurvfit(fit_OS, size = 1.5) +
    add_censor_mark() +
    add_confidence_interval()+
    scale_ggsurvfit() +
    add_risktable(
      size            = 7,
      theme           = theme_risktable_default(axis.text.y.size = 10,
                                                 plot.title.size  = 20),
      risktable_stats = "{n.risk}",  # ({cum.event}) removed because of space
      stats_label = "Number at risk"
    ) +
    add_risktable_strata_symbol(symbol = "•", size = 20)+
    labs(
      x        = "Months after treatment initiation",
      y        = "PFS (%)",
      title = "RCC - Overall Survival"
    ) +
    scale_x_continuous(breaks = seq(0, x_max, by = 12), limits = c(0, x_max)) +
    theme_classic() +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 25),
      axis.title.x    = element_text(size = 20),
      axis.title.y    = element_text(size = 20, margin = margin(t = 0, r = 0, b = 0, l = 0)),
      axis.text.x     = element_text(size = 15),
      axis.text.y     = element_text(size = 15),
      legend.position = "bottom",
      legend.direction= "horizontal",
      legend.text     = element_text(size = 18),
      legend.key.size = unit(10, "bigpts"),
      legend.title    = element_blank(),
      plot.margin = unit(c(0,0.2,0,1), 'lines')
    ) +
    scale_color_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
    scale_fill_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
   annotate(
    "text",
    x     = x_max*0.25,
    y     = ypos[names(ypos)]*0.9/100,
    label = paste0("5 year OS for ",os_lbls$os_lbls),
    size  = 10,
    hjust = 0
    )


# PFS plot
ypos <- c(CR = five_year_table$PFS_5yr_pct[1], PR = five_year_table$PFS_5yr_pct[2])

fit_PFS <- survfit(
  Surv(censor_time_PFS, censor_status_PFS) ~ Objective_response,
  data = dataDF
)
names(fit_PFS$strata) <- gsub("Objective_response=", "", names(fit_PFS$strata))

p_pfs <- ggsurvfit(fit_PFS, size = 1.5) +
    add_censor_mark() +
    add_confidence_interval()+
    scale_ggsurvfit() +
    add_risktable(
      size            = 7,
      theme           = theme_risktable_default(axis.text.y.size = 10,
                                                 plot.title.size  = 20),
      risktable_stats = "{n.risk}",  # ({cum.event}) removed because of space
      stats_label = "Number at risk"
    ) +
    add_risktable_strata_symbol(symbol = "•", size = 20)+
    labs(
      x        = "Months after treatment initiation",
      y        = "PFS (%)",
      title = "RCC - Progression-free Survival"
    ) +
    scale_x_continuous(breaks = seq(0, x_max, by = 12), limits = c(0, x_max)) +
    theme_classic() +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 25),
      axis.title.x    = element_text(size = 20),
      axis.title.y    = element_text(size = 20, margin = margin(t = 0, r = 0, b = 0, l = 0)),
      axis.text.x     = element_text(size = 15),
      axis.text.y     = element_text(size = 15),
      legend.position = "bottom",
      legend.direction= "horizontal",
      legend.text     = element_text(size = 18),
      legend.key.size = unit(10, "bigpts"),
      legend.title    = element_blank(),
      plot.margin = unit(c(0,0.2,0,1), 'lines')
    ) +
    scale_color_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
    scale_fill_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
   annotate(
    "text",
    x     = x_max*0.25,
    y     = ypos[names(ypos)]*0.9/100,
    label = paste0("5 year PFS for ",pfs_lbls$pfs_lbls),
    size  = 10,
    hjust = 0
    )

p_pfs
p_os
```


## 18 Month landmark analysis
```{r landmark analysis, warning=FALSE, message=FALSE, fig.width=10, fig.height=8}
#remove all patients with an event before 18 months
dataDF_18mo <- subset(dataDF, OS_days > 540 & PFS_days > 540)

#reset time
dataDF_18mo$censor_time_OS <- dataDF_18mo$censor_time_OS-18
dataDF_18mo$censor_time_PFS <- dataDF_18mo$censor_time_PFS-18

surv_time <- "censor_time_PFS"
surv_status <- "censor_status_PFS"
surv_time_label <- "PFS"
# PFS plot
results <- lapply(responses, function(resp) {
  d <- filter(dataDF_18mo, `Objective response` == resp)
  fit_OS  <- survfit(Surv(censor_time_OS,  censor_status_OS)  ~ 1, data = d)
  fit_PFS <- survfit(Surv(censor_time_PFS, censor_status_PFS) ~ 1, data = d)

  os_ci  <- get5yr_ci(fit_OS)
  pfs_ci <- get5yr_ci(fit_PFS)


  data.frame(
    Response       = resp,
    OS_5yr         = os_ci$Estimate,
    OS_5yr_Lower   = os_ci$Lower95,
    OS_5yr_Upper   = os_ci$Upper95,
    PFS_5yr        = pfs_ci$Estimate,
    PFS_5yr_Lower  = pfs_ci$Lower95,
    PFS_5yr_Upper  = pfs_ci$Upper95
  )
})

five_year_table <- bind_rows(results) %>%
  mutate(across(matches("_5yr"), ~ . * 100, .names = "{.col}_pct")) %>%
  select(Response,
         OS_5yr_pct, OS_5yr_Lower_pct, OS_5yr_Upper_pct,
         PFS_5yr_pct, PFS_5yr_Lower_pct, PFS_5yr_Upper_pct,
         )


pfs_lbls <- five_year_table %>%
  transmute(
    Response,
    pfs_label = sprintf("%s: %.0f%% (%.0f–%.0f%%)",
                        Response, PFS_5yr_pct, PFS_5yr_Lower_pct, PFS_5yr_Upper_pct))

ypos <- c(CR = five_year_table$PFS_5yr_pct[1], PR = five_year_table$PFS_5yr_pct[2])

fit_PFS <- survfit(
  Surv(censor_time_PFS, censor_status_PFS) ~ Objective_response,
  data = dataDF_18mo
)
names(fit_PFS$strata) <- gsub("Objective_response=", "", names(fit_PFS$strata))

x_max <- 32
p_pfs <- ggsurvfit(fit_PFS, size = 1.5) +
    add_censor_mark() +
    add_confidence_interval()+
    scale_ggsurvfit() +
    add_risktable(
      size            = 7,
      theme           = theme_risktable_default(axis.text.y.size = 10,
                                                 plot.title.size  = 20),
      risktable_stats = "{n.risk}",  # ({cum.event}) removed because of space
      stats_label = "Number at risk"
    ) +
    add_risktable_strata_symbol(symbol = "•", size = 20)+
    labs(
      x        = "Months after treatment initiation",
      y        = "PFS (%)",
      title = "RCC - Progression-free Survival"
    ) +
    scale_x_continuous(breaks = seq(0, x_max, by = 6), limits = c(0, x_max)) +
    theme_classic() +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 25),
      axis.title.x    = element_text(size = 20),
      axis.title.y    = element_text(size = 20, margin = margin(t = 0, r = 0, b = 0, l = 0)),
      axis.text.x     = element_text(size = 15),
      axis.text.y     = element_text(size = 15),
      legend.position = "bottom",
      legend.direction= "horizontal",
      legend.text     = element_text(size = 18),
      legend.key.size = unit(10, "bigpts"),
      legend.title    = element_blank(),
      plot.margin = unit(c(0,0.2,0,1), 'lines')
    ) +
    scale_color_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
    scale_fill_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
   annotate(
    "text",
    x     = x_max*0.25,
    y     = (ypos[names(ypos)]*0.9)/100,
    label = paste0("5 year PFS for ",pfs_lbls$pfs_label),
    size  = 10,
    hjust = 0
    )

os_lbls <- five_year_table %>%
  transmute(
    Response,
    os_label = sprintf("%s: %.0f%% (%.0f–%.0f%%)",
                        Response, OS_5yr_pct, OS_5yr_Lower_pct, OS_5yr_Upper_pct))

ypos <- c(CR = five_year_table$OS_5yr_pct[1], PR = five_year_table$OS_5yr_pct[2])

fit_OS <- survfit(
  Surv(censor_time_OS, censor_status_OS) ~ Objective_response,
  data = dataDF_18mo
)
names(fit_OS$strata) <- gsub("Objective_response=", "", names(fit_OS$strata))

p_os <- ggsurvfit(fit_OS, size = 1.5) +
    add_censor_mark() +
    add_confidence_interval()+
    scale_ggsurvfit() +
    add_risktable(
      size            = 7,
      theme           = theme_risktable_default(axis.text.y.size = 10,
                                                 plot.title.size  = 20),
      risktable_stats = "{n.risk}",  # ({cum.event}) removed because of space
      stats_label = "Number at risk"
    ) +
    add_risktable_strata_symbol(symbol = "•", size = 20)+
    labs(
      x        = "Months after treatment initiation",
      y        = "OS (%)",
      title = "RCC - Overall Survival"
    ) +
    scale_x_continuous(breaks = seq(0, x_max, by = 6), limits = c(0, x_max)) +
    theme_classic() +
    theme(
      plot.title      = element_text(hjust = 0.5, size = 25),
      axis.title.x    = element_text(size = 20),
      axis.title.y    = element_text(size = 20, margin = margin(t = 0, r = 0, b = 0, l = 0)),
      axis.text.x     = element_text(size = 15),
      axis.text.y     = element_text(size = 15),
      legend.position = "bottom",
      legend.direction= "horizontal",
      legend.text     = element_text(size = 18),
      legend.key.size = unit(10, "bigpts"),
      legend.title    = element_blank(),
      plot.margin = unit(c(0,0.2,0,1), 'lines')
    ) +
    scale_color_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
    scale_fill_manual(name=c("CR", "PR"),values = c("CR"="#619CFF", "PR"="#F564E3")) +
   annotate(
    "text",
    x     = x_max*0.25,
    y     = (ypos[names(ypos)]*0.9)/100,
    label = paste0("5 year OS for ",os_lbls$os_label),
    size  = 10,
    hjust = 0
    )

p_pfs
p_os
```
